Numerical simulation of the Cordilleran ice sheet \\ through the last glacial cycle

### 2014-06-25 kappa.tex
Add paper-n commands as the order is unsure.

@@

The till cohesion $c_0$ is assumed to be zero. The effective pressure on the
till, $N$, is determined by the modelled amount of water at the bed. However,
two different parametrisations are used within this thesis. In Paper~I, using
PISM~0.5, the effective pressure is linearly related to water content by a
simple parametrisation,
\begin{equation}
    N = \rho gH (1 - \alpha \frac{W}{W_{max}}) \,.
\end{equation}

In Paper~IV, using PISM~0.6, the effective pressure is determined by an
empirical parametrisation based on laboratory experiments on Antarctic till,
\begin{equation}
    N = \delta \rho gH \, 10^{(e_0/C_c) (1 - (W/W_{max}))} \,.
\end{equation}

@@

Finally, the till friction angle $\phi$ is taken a~function of modern bed
elevation, with lowest values occurring at low elevations, thus accounting for
a weakening of till associated with the presence of marine sediments:
\begin{equation}
    \phi(x,y) =
    \begin{cases}
        \phi_0 & \text{if}\ b \le b_0 \,, \\
        \phi_0 + (\phi_1-\phi_0) \frac{b - b_0}{b_1-b_0}
                & \text{if}\ b_0 < b < b_1 \,, \\
        \phi_1 & \text{if}\ b_1 \le b \,,
    \end{cases}
\end{equation}
where $b_0=0$\,m is modern sea-level and $b_1=200$\,m is a rough average of
highest shoreline elevations in the model domain. The values of $\phi_1$ and
$\phi_2$ are chosen differently in Paper~I (10 and 30$^\circ$) and Paper~IV
(15 and 45$^\circ$), but these choices are roughly equivalent in the case of a
saturated till ($W=W_{max}$).

@@

In the simulations presented here, \cref{eqn:calovgreve} is numerically
approximated using week-long sub-intervals. For each sub-interval, surface melt
is computed from PDDs and the available snow and ice depths using an algorithm
documented in the on-line documentation of PISM \citep{PISM-authors.2014}. This
algorythms employ different melt factors for snow ($F_s$) and for ice ($F_i$),
as derived from mass-balance measurements on contemporary glaciers in the
Coast and Rocky Mountains of British Columbia \citep{Shea.etal.2009}. In
Paper~I, the temperature standard deviation $\sPDD$ is a constant model
parameter. In Paper~IV, it is computed from daily temperature values from the
North American Regional Reanalysis \citep[NARR,][]{Mesinger.etal.2006}, using
a method initially developed in Paper~II, and further improved in Paper~III.
%3.04\,\unit{mm\,{\degree}C^{-1}\,day^{-1}} for snow and
%4.59\,\unit{mm\,{\degree}C^{-1}\,day^{-1}} for ice

@@

Paper~I tests the sensitivity of the model to different input present-day
climatologies, $\{T_{m0}, P_{m0}\}$, while the temperature-offset corrections,
${\Delta}T_{TS}$, uses constant values ranging from -10 to 0\,\degC. Paper~IV,
on the opposite, tests model sensitivity to time-dependent temperature-offset
corrections derived from different palaeo-temperature proxy records.

>>>>

% Papers of the thesis
\newcommand{\CCLI}[0]{Paper~I}      % Cordillera climate
\newcommand{\PSDV}[0]{Paper~II}     % PDD SD variability
\newcommand{\PSDP}[0]{Paper~III}    % PDD SD parametrisation
\newcommand{\CCYC}[0]{Paper~IV}     % Cordillera cycle

@@

The till cohesion $c_0$ is assumed to be zero. The effective pressure on the
till, $N$, is determined by the modelled amount of water at the bed. However,
two different parametrisations are used within this thesis. In {\CCLI}, using
PISM~0.5, the effective pressure is linearly related to water content by a
simple parametrisation,
\begin{equation}
    N = \rho gH (1 - \alpha \frac{W}{W_{max}}) \,.
\end{equation}

In {\CCYC}, using PISM~0.6, the effective pressure is determined by an
empirical parametrisation based on laboratory experiments on Antarctic till,
\begin{equation}
    N = \delta \rho gH \, 10^{(e_0/C_c) (1 - (W/W_{max}))} \,.
\end{equation}

@@

Finally, the till friction angle $\phi$ is taken a~function of modern bed
elevation, with lowest values occurring at low elevations, thus accounting for
a weakening of till associated with the presence of marine sediments:
\begin{equation}
    \phi(x,y) =
    \begin{cases}
        \phi_0 & \text{if}\ b \le b_0 \,, \\
        \phi_0 + (\phi_1-\phi_0) \frac{b - b_0}{b_1-b_0}
                & \text{if}\ b_0 < b < b_1 \,, \\
        \phi_1 & \text{if}\ b_1 \le b \,,
    \end{cases}
\end{equation}
where $b_0=0$\,m is modern sea-level and $b_1=200$\,m is a rough average of
highest shoreline elevations in the model domain. The values of $\phi_1$ and
$\phi_2$ are chosen differently in \CCLI (10 and 30$^\circ$) and \CCYC
(15 and 45$^\circ$), but these choices are roughly equivalent in the case of a
saturated till ($W=W_{max}$).

@@

In the simulations presented here, \cref{eqn:calovgreve} is numerically
approximated using week-long sub-intervals. For each sub-interval, surface melt
is computed from PDDs and the available snow and ice depths using an algorithm
documented in the on-line documentation of PISM \citep{PISM-authors.2014}. This
algorythms employ different melt factors for snow ($F_s$) and for ice ($F_i$),
as derived from mass-balance measurements on contemporary glaciers in the
Coast and Rocky Mountains of British Columbia \citep{Shea.etal.2009}. In
{\CCLI}, the temperature standard deviation $\sPDD$ is a constant model
parameter. In {\CCYC}, it is computed from daily temperature values from the
North American Regional Reanalysis \citep[NARR,][]{Mesinger.etal.2006}, using
a method initially developed in {\PSDV}, and further improved in {\PSDP}.
%3.04\,\unit{mm\,{\degree}C^{-1}\,day^{-1}} for snow and
%4.59\,\unit{mm\,{\degree}C^{-1}\,day^{-1}} for ice

@@

\CCLI tests the sensitivity of the model to different input present-day
climatologies, $\{T_{m0}, P_{m0}\}$, while the temperature-offset corrections,
${\Delta}T_{TS}$, uses constant values ranging from -10 to 0\,\degC. \CCYC,
on the opposite, tests model sensitivity to time-dependent temperature-offset
corrections derived from different palaeo-temperature proxy records.
